---
title: "Horseweed emergence in Nebraska"
author: "Maxwel Coura Oliveira"
date: "6/29/2021"
output:       
  html_document:  
    css: "my_css.css" 
    toc: TRUE
    
header-includes:
  - \setcounter{section}{4}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 7,
  fig.width = 7,
  fig.align = 'center')
```


```{r}
options(scipen = 999)
library(tidyverse)
library(drc)
library(reactable)
library(reactablefmtr)
```

I have used the same model for all locations. You will see some comments in each subsection.

## Scottsbluff

```{r}
readxl::read_excel("Emergence.xlsx", sheet = "SCOTTSBLUFF") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  dplyr::select(1:4) %>% 
  pivot_longer(cols = 2:4, names_to = "type", values_to = "emergence") %>% 
  mutate(date = "Sep 2016 - Apr 2017") %>% 
  mutate_at(c("GDD", "emergence"),  as.double) %>% 
  mutate_at(c("GDD", "emergence"), ~ round(., 2)) %>% 
  separate("type", c("population", "type"), sep = "-") %>% 
  mutate_at(c("population", "type"), str_squish) %>% 
  mutate(population = fct_recode(population,
                                 "Lin1" = "Lincoln",
                                 "Npl1" = "North Platte",
                                 "Scb1" = "Scottsbluff")) %>% 
  mutate(location = "Scottsbluff") %>% 
  filter(!is.na(GDD)) -> sctbf1
```

```{r}
readxl::read_excel("Emergence.xlsx", sheet = "SCOTTSBLUFF") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  dplyr::select(7:13) %>% 
  pivot_longer(cols = 2:7, names_to = "type", values_to = "emergence") %>% 
  mutate(date = "Sep 2017 - Dec 2017") %>% 
  mutate_at(c("GDD", "emergence"),  as.double) %>% 
  mutate_at(c("GDD", "emergence"), ~ round(., 2)) %>% 
  separate("type", c("population", "type"), sep = "-") %>% 
  mutate_at(c("population", "type"), str_squish) %>% 
  mutate(population = fct_recode(population,
                                 "Lin1" = "Lincoln",
                                 "Npl1" = "North Platte",
                                 "Scb1" = "Scottsbluff")) %>% 
  mutate(location = "Scottsbluff") -> sctbf2
```


```{r}
sctbf1 %>% 
  bind_rows(sctbf2) %>% 
  mutate(type = str_to_lower(type)) %>% 
  unite("trt", c("population", "type"), sep = " in ", remove = FALSE) %>% 
  drop_na() -> scottsbluff
```

### Model

I am using a Weibull function with three parameters. Notice that the upper limit is locked at 100.


```{r echo = TRUE}
model1 <- drm(emergence ~ GDD, trt, 
              fct = W1.3(fixed = c(NA, 100, NA), 
                         names = c("slope", "upper", "ed50")),
              data = scottsbluff)
```

### Figure

Do not worried about the figure right now because I am uging `ggplot` to make a better viz.

```{r echo = TRUE}
plot(model1, legendPos = c(40, 95), col = c(1,2,3,4,5,6),
     ylab = "Emergence (%)")
```

### Model parameters

```{r include=FALSE}
broom::tidy(model1) %>% 
  mutate_if(is.double, ~ round(., 4)) -> param_sct
```

```r
summary(model1)
```

```{r}
param_sct %>% 
  reactable(theme = fivethirtyeight())
```


### EDs

ED comes from effective dose; however we are using the model for weed emergence. In this case, each 10, 50, and 90 represent the % of emergence for each treatment.

```{r include=FALSE}
ED(model1, c(10, 50, 90), interval = "delta") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Term") %>% 
  mutate_if(is.double, ~ round(., 4)) -> ed_sct
```

```r
ED(model2, c(10, 50, 90), interval = "delta")
```

```{r}
ed_sct %>% 
  reactable(theme = fivethirtyeight()) 
```



### Emergence comparisons

A t-test is made for each treatment at 90%, 50%, and 10%. You can compare 10 with 90% also. Also, you can compare any number between 1 and 100. I am using 10, 50, and 90% because it is the most common and you can tell a nice story.

You can use **p-values** to conclude if a treatment is different to another. 

#### Comparison of 90% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model1, c(90,90), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcompsct90
```

```r
EDcomp(model1, c(90,90), type = "absolute")
```

```{r}
edcompsct90 %>% 
  reactable(theme = fivethirtyeight())
```

#### Comparison of 50% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model1, c(50,50), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcompsct50
```

```r
EDcomp(model1, c(50,50), type = "absolute")
```

```{r}
edcompsct50 %>% 
  reactable(theme = fivethirtyeight())
```



#### Comparison of 10% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model1, c(10,10), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcompsct10
```

```r
EDcomp(model1, c(10,10), type = "absolute")
```

```{r}
edcompsct10 %>% 
  reactable(theme = fivethirtyeight())
```



## Lincoln


```{r}
readxl::read_excel("Emergence.xlsx", sheet = "Lincoln") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  dplyr::select(1:9) %>% 
  pivot_longer(cols = 2:9, names_to = "type", values_to = "emergence") %>% 
  mutate(date = "Sep 2016 - Apr 2017") %>% 
  mutate_at(c("GDD", "emergence"),  as.double) %>% 
  mutate_at(c("GDD", "emergence"), ~ round(., 2)) %>% 
  separate("type", c("population", "type"), sep = "-") %>% 
  mutate_at(c("population", "type"), str_squish) %>% 
  mutate(population = fct_recode(population,
                                 "Lin1" = "Lincoln",
                                 "Npl1" = "North Platte",
                                 "Scb1" = "Scottsbluff")) %>% 
  mutate(location = "Lincoln") %>% 
  filter(!is.na(GDD)) -> lcn1
```


```{r}
readxl::read_excel("Emergence.xlsx", sheet = "Lincoln") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  dplyr::select(13:21) %>% 
  pivot_longer(cols = 2:9, names_to = "type", values_to = "emergence") %>% 
  mutate(date = "Sep 2017-Jun 2018") %>% 
  mutate_at(c("GDD", "emergence"),  as.double) %>% 
  mutate_at(c("GDD", "emergence"), ~ round(., 2)) %>% 
  separate("type", c("population", "type"), sep = "-") %>% 
  mutate_at(c("population", "type"), str_squish) %>% 
  mutate(population = fct_recode(population,
                                 "Lin1" = "Lincoln",
                                 "Npl1" = "North Platte",
                                 "Scb1" = "Scottsbluff")) %>% 
  mutate(location = "Lincoln") %>% 
  filter(!is.na(GDD)) -> lcn2
```


```{r}
lcn1 %>% 
  bind_rows(lcn2) %>% 
  mutate(type = str_to_lower(type)) %>% 
  unite("trt", c("population", "type"), sep = " in ", remove = FALSE) %>% 
  drop_na() -> lincoln
```

### Model

I am using a Weibull function with three parameters. Notice that the upper limit is locked at 100.

I have used the same model for all locations. 

```{r echo = TRUE}
model2 <- drm(emergence ~ GDD, trt, 
              fct = W1.3(fixed = c(NA, 100, NA), 
                         names = c("slope", "upper", "ed50")),
              data = lincoln)
```

### Figure

```{r}
plot(model2, legendPos = c(70, 95), col = c(1,2,3,4,5,6),
     ylab = "Emergence (%)")
```

### Model parameters

```{r include=FALSE}
broom::tidy(model2) %>% 
  mutate_if(is.double, ~ round(., 4)) -> param_lin
```

```r
summary(model2)
```

```{r}
param_lin %>% 
  reactable(theme = fivethirtyeight())
```


### EDs

```{r include=FALSE}
ED(model2, c(10, 50, 90), interval = "delta") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Term") %>% 
  mutate_if(is.double, ~ round(., 4)) -> ed_lin
```


```r
ED(model2, c(10, 50, 90), interval = "delta")
```


```{r}
ed_lin %>% 
  reactable(theme = fivethirtyeight())
```


### Emergence comparisons

A t-test is made for each treatment at 90%, 50%, and 10%. You can compare 10 with 90% also. Also, you can compare any number between 1 and 100. I am using 10, 50, and 90% because it is the most common and you can tell a nice story.

#### Comparison of 90% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model2, c(90,90), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcomplin90
```

```r
EDcomp(model2, c(90,90), type = "absolute")
```

```{r}
edcomplin90 %>% 
  reactable(theme = fivethirtyeight())
```

#### Comparison of 50% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model2, c(50,50), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcomplin50
```

```r
EDcomp(model2, c(50,50), type = "absolute")
```

```{r}
edcomplin50 %>% 
  reactable(theme = fivethirtyeight())
```



#### Comparison of 10% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model2, c(10,10), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcomplin10
```

```r
EDcomp(model2, c(10,10), type = "absolute")
```

```{r}
edcomplin10 %>% 
  reactable(theme = fivethirtyeight())
```



## North Platte

```{r}
readxl::read_excel("Emergence.xlsx", sheet = "North Platte") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  dplyr::select(1:4) %>% 
  pivot_longer(cols = 2:4, names_to = "type", values_to = "emergence") %>% 
  mutate(date = "Sep 2016 - May 2017") %>% 
  mutate_at(c("GDD", "emergence"),  as.double) %>% 
  mutate_at(c("GDD", "emergence"), ~ round(., 2)) %>% 
  separate("type", c("population", "type"), sep = "-") %>% 
  mutate_at(c("population", "type"), str_squish) %>% 
  mutate(population = fct_recode(population,
                                 "Lin1" = "Lincoln",
                                 "Npl1" = "North Platte",
                                 "Scb1" = "Scottsbluff")) %>% 
  mutate(location = "North Platte") %>% 
  filter(!is.na(GDD)) -> np1
```


```{r}
readxl::read_excel("Emergence.xlsx", sheet = "North Platte") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  dplyr::select(5:8) %>% 
  pivot_longer(cols = 2:4, names_to = "type", values_to = "emergence") %>% 
  mutate(date = "Sep 2016 - May 2017") %>% 
  mutate_at(c("GDD", "emergence"),  as.double) %>% 
  mutate_at(c("GDD", "emergence"), ~ round(., 2)) %>% 
  separate("type", c("population", "type"), sep = "-") %>% 
  mutate_at(c("population", "type"), str_squish) %>% 
  mutate(population = fct_recode(population,
                                 "Lin1" = "Lincoln",
                                 "Npl1" = "North Platte",
                                 "Scb1" = "Scottsbluff")) %>% 
  mutate(location = "North Platte") %>% 
  filter(!is.na(GDD)) -> np2
```



```{r}
np1 %>% 
  bind_rows(np2) %>% 
  mutate(type = str_to_lower(type)) %>% 
  unite("trt", c("population", "type"), sep = " in ", remove = FALSE) %>% 
  drop_na() -> north_platte
```


### Model

```{r echo = TRUE}
model3 <- drm(emergence ~ GDD, trt, 
              fct = W1.3(fixed = c(NA, 100, NA), 
                         names = c("slope", "upper", "ed50")),
              data = north_platte)
```

### Figure 


```{r}
plot(model3, legendPos = c(70, 95), col = c(1,2,3,4,5,6),
     ylab = "Emergence (%)")
```





### Model parameters

```{r include=FALSE}
broom::tidy(model3) %>% 
  mutate_if(is.double, ~ round(., 4)) -> param_np
```

```r
summary(model2)
```

```{r}
param_np %>% 
  reactable(theme = fivethirtyeight())
```



### EDs

```{r include=FALSE}
ED(model3, c(10, 50, 90), interval = "delta") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Term") %>% 
  mutate_if(is.double, ~ round(., 4)) -> ed_np
```

```r
ED(model3, c(10, 50, 90), interval = "delta")
```

```{r}
ed_np %>% 
  reactable(theme = fivethirtyeight())
```


### Emergence comparisons

A t-test is made for each treatment at 90%, 50%, and 10%. You can compare 10 with 90% also. Also, you can compare any number between 1 and 100. I am using 10, 50, and 90% because it is the most common and you can tell a nice story.

#### Comparison of 90% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model3, c(90,90), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcompnp90
```

```r
EDcomp(model3, c(90,90), type = "absolute")
```

```{r}
edcompnp90 %>% 
  reactable(theme = fivethirtyeight())
```

#### Comparison of 50% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model3, c(50,50), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcompnp50
```

```r
EDcomp(model3, c(50,50), type = "absolute")
```

```{r}
edcompnp50 %>% 
  reactable(theme = fivethirtyeight())
```



#### Comparison of 10% emergence

```{r echo=TRUE, include=FALSE}
EDcomp(model3, c(10,10), type = "absolute") %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "Comparison") %>% 
  mutate_if(is.double, ~ round(., 4)) -> edcompnp10
```

```r
EDcomp(model3, c(10,10), type = "absolute")
```

```{r}
edcompnp10 %>% 
  reactable(theme = fivethirtyeight())
```